---
title: "Homework 3"
author: "Yuki Joyama"
date: "2023-10-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F)
```

```{r}
library(tidyverse)

# import data set from {MASS} package
library(MASS)
df = birthwt
```

# Problem 1
```{r}
n = length(pull(df, lwt))
x = mean(pull(df, lwt))
s = sd(pull(df, lwt))

lower_ci = x - qt(.975, df = n-1)*(s/sqrt(n))
upper_ci = x + qt(.975, df = n-1)*(s/sqrt(n))
```
Let n = `r length(pull(df, lwt))`, $\bar{X}$ (sample mean) = `r round(mean(pull(df, lwt)), 2)`, $s$ (sample standard deviation) = `r round(sd(pull(df, lwt)), 2)`

a) Because we do not know the true population variance $\sigma^2$, 95% confidence interval of true mean weight can be obtained by the following:  
$$\bar{X}-t_{n-1,1-\alpha/2}\frac{s}{\sqrt{n}}, \bar{X}+t_{n-1,1-\alpha/2}\frac{s}{\sqrt{n}}$$

Thus, 95% confidence interval is (`r round(lower_ci, 2)`, `r round(upper_ci, 2)`)

b) We are 95% confident that the true mean weight of American women will be between (`r round(lower_ci, 2)`, `r round(upper_ci, 2)`).

c) The statement is not valid because 171 pounds is not included in the 95% confidence interval.

# Problem 2
```{r results=F}
# equality of variances between smoking and non-smoking group (alpha = 0.05)
s1 = df |> 
  filter(smoke == 0) |> 
  summarise(sd = sd(lwt))
  
s2 = df |> 
  filter(smoke == 1) |> 
  summarise(sd = sd(lwt))

n1 = df |> 
  filter(smoke == 0) |> 
  drop_na(lwt) |> 
  nrow()

n2 = df |> 
  filter(smoke == 1) |> 
  drop_na(lwt) |> 
  nrow()

# check
var.test(lwt ~ smoke, df, 
         alternative = "two.sided")

# two sample independent t-test for unequal variances
m1 = df |> 
  filter(smoke == 0) |> 
  summarise(mean = mean(lwt))

m2 = df |> 
  filter(smoke == 1) |> 
  summarise(mean = mean(lwt))

pool_sd = sqrt(((n1-1)*s1^2+(n2-1)*s2^2)/(n1+n2-2))

# check
t.test(lwt ~ smoke, var.equal = TRUE, df)
```

a) Let $s_1$ and $s_2$ be the sample standard deviation from non-smoking group and smoking group accordingly. And let $n_1$ and $n_2$ be the number of observations in each group.  
$\alpha=0.05$  
$H_0$: the variances in two groups are equal  
$H_1$: the variances in two groups are not equal  

The test statistic is  
$F = \frac{s_1^2}{s_2^2}=$ `r round((s1^2)/(s2^2), 2)` ~ $F_{n_1-1, n_2-1}$  

Now, $F_{n_1-1, n_2-1, 1-\alpha/2}=$ `r round(qf(.975, n1-1, n2-1), 2)` > F  
Thus, we fail to reject the null hypothesis and accept $H_0$.   

b) The answer from part a indicates that the variances between two groups are equal at 5% significance level. Therefore, I will perform two-sample independent t-test for equal variances. 

c) $\alpha=0.10$ 
Let $\mu_1$ and $\mu_2$ be the population mean of the two groups. ($\bar{X_1}, \bar{X_2}$: sample mean)  
$H_0$: $\mu_1=\mu_2$  
$H_1$: $\mu_1\neq\mu_2$  

The pooled estimate of the standard deviation $s = \sqrt{\frac{(n_1-1)s_1^2+ (n_2-1)s_2^2}{n_1+n_2-2}}=$ `r round(pool_sd, 2)`  

The test statistic is  
$t=\frac{\bar{X_1}- \bar{X_2}}{s\sqrt{\frac{1}{n_1}+\frac{1}{n_2}}}=$ `r round((m1-m2)/(pool_sd*sqrt(1/n1+1/n2)), 2)` ~ $t_{n_1+n_2-2}$  

$t_{n_1+n_2-2,1-\alpha/2}=$ `r round(qt(.95, n1 + n2 -2), 2)` > t so we fail to reject the null hypothesis and accept $H_0$.   
At 10% significance level, we can say that there is no statistically significant difference in the mother's weight at last menstrual period between smoking and non-smoking groups.

# Problem 3
```{r}
# 99% CI
df3 = df |> 
  group_by(ht) |> 
  summarise(n = n()) |> 
  mutate(total = sum(n))

p = df3 |> 
  filter(ht == 1) |> 
  mutate(p = n / total) |> 
  pull(p)

p0 = 0.2

n = df |> 
  drop_na(ht) |> 
  nrow()

lower_ci = p - qnorm(.995) * sqrt(p*(1-p)/n)
upper_ci = p + qnorm(.995) * sqrt(p*(1-p)/n)

z = (p-p0)/sqrt(p0*(1-p0)/n)
```

a) $\alpha=0.01$  
Let p be the proportion of pregnant American women suffer from hypertension. ($p_0=$ 0.20)  
I will perform one-sample test for binomial proportion.  

99% confidence interval for one population proportion can be calculated by:  
$$
\hat{p}-z_{1-\alpha/2}\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}, \hat{p}+z_{1-\alpha/2}\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}
$$

Thus, 99% CI is (`r round(lower_ci, 2)`, `r round(upper_ci, 2)`)

Because 0.20 is not included in the 99% CI, we can conclude that we are 99% confident that our data does not support this claim. 

b) $\alpha=0.01$  
$H_0$: $p=p_0$    
$H_1$: $p <p_0$   

Test statistic is $z=\frac{\hat{p}-p_0}{\sqrt{p_0(1-p_0)/n}}=$ `r round(z, 2)` ~ N(0, 1)

$z_{\alpha}=$ `r round(qnorm(.1), 2)` > $z$  
Therefore, we reject the null hypothesis and accept $H_1$. 
We can conclude that at 10% significant level, the true proportion is indeed less than the claimed 20%. 
# Problem 4







