---
title: "Homework 3"
author: "Yuki Joyama"
date: "2023-10-23"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F)
```

```{r}
library(tidyverse)

# import data set from {MASS} package
library(MASS)
df = birthwt
```

# Problem 1
```{r}
n = length(pull(df, lwt))
x = mean(pull(df, lwt))
s = sd(pull(df, lwt))

lower_ci = x - qt(.975, df = n-1)*(s/sqrt(n))
upper_ci = x + qt(.975, df = n-1)*(s/sqrt(n))
```
Let n = `r length(pull(df, lwt))`, $\bar{X}$ (sample mean) = `r round(mean(pull(df, lwt)), 2)`, $s$ (sample standard deviation) = `r round(sd(pull(df, lwt)), 2)`

a) Because we do not know the true population variance $\sigma^2$, 95% confidence interval of true mean weight can be obtained by the following:  
$$\bar{X}-t_{n-1,1-\alpha/2}\frac{s}{\sqrt{n}}, \bar{X}+t_{n-1,1-\alpha/2}\frac{s}{\sqrt{n}}$$

Thus, 95% confidence interval is (`r round(lower_ci, 2)`, `r round(upper_ci, 2)`)

b) We are 95% confident that the true mean weight of American women will be between (`r round(lower_ci, 2)`, `r round(upper_ci, 2)`).

c) The statement is not valid because 171 pounds is not included in the 95% confidence interval.

# Problem 2








